There once was a grid at ol' Carkeek

### 2016-08-13 Manuscript/v0.0/carkeek_grid.tex
add PCR methods

>>>>

\subsection*{PCR Amplification}

From each DNA sample, we amplified an approximately 115 base pair (bp) region of the mitochondrial gene encoding 16S RNA using a two-step polymerase chain reaction (PCR) protocol described by ODONNELL (CITE ODONNELL2016).
In the first step, primers were identical in every reaction (forward: AGTTACYYTAGGGATAACAGCG; reverse: CCGGTCTGAACTCAGATCAYGT); primers in the second step included these same sequences but with 3 variable nucleotides ('NNN') and an index sequence on the 5PRIME end (see CITE PRIMER TABLE/METADATA).
We used the program OligoTag \cite{Coissac2012} to generate 30 unique 6 nucleotide index sequences differing by a minimum Hamming distance of 3 (\nameref{primer_table}).
Indexed primers were assigned to samples randomly, with the identical index sequence on the forward and reverse primer to avoid errors associated with dual-indexed multiplexing (CITE Schnell2015).
In a UV-sterilized hood, we prepared 25 microliter reactions containing 18.375 microliters ultrapure water, 2.5 microliters 10x buffer, 0.625 microliters deoxynucleotide solution (8 millimolar), 1 microliter each forward and reverse primer (10 micromolar, obtained lyophilized from Integrated DNA Technologies (Coralville, IA, USA)), 0.25 microliters Qiagen HotStar Taq polymerase, and 1.25 microliter genomic eDNA template at 1:100 dilution in ultrapure water.
PCR thermal profiles began with an initialization step (95C; 15 min) followed by cycles (40 and 20 for the first and second reaction, respectively) of denaturation (95C; 15 sec), annealing (61C; 30 sec), and extension (72C; 30 sec).
20 identical PCRs were conducted from each DNA extract using non-indexed primers; these were pooled into 4 groups of 5 in order to ensure ample template for the subsequent PCR with indexed primers. % we probably pooled >1 product from PCR1, but figuring this out from our lab notebook would be time consuming and perhaps impossible.
In order to isolate the fragment of interest from primer dimer and other spurious fragments generated in the first PCR, we used the AxyPrep Mag FragmentSelect-I kit with solid-phase reversible immobilization (SPRI) paramagnetic beads at 2.5x the volume of PCR product (Axygen BioSciences, Corning, NY, USA).
A 1:5 dilution in ultrapure water of the product was used as template for the second reaction.
PCR products of the second reaction were purified using the Qiagen MinElute PCR Purification Kit (Qiagen, Hilden, Germany).
Ultrapure water was used in place of template DNA and run along with each batch of PCRs to serve as a negative control for PCR; none of these produced visible bands on an agarose gel.
In total, four separate replicates from each of 31 DNA samples were carried through the two-step PCR process for a total of 124 sequenced PCR products.
These were combined with additional samples from other projects, totaling 345 samples.
